---
title: "NEID Solar Data -- EDA"
author: "Joe Salzer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("readSpectra.R")
library(collapse)
library(patchwork)
```

# Loading/cleaning data

variables to be set by the user

```{r}
# get csv file name
csvFileName = "completeLines.csv"
```

```{r}
# working directory with the data
WD_DATA = "/Users/josephsalzer/research/exostat/"
# names of the timeID, lineID, and timeGroupID
TIME_ID_NAME = "date"
LINE_ID_NAME = "line_order"
TIMEGROUP_ID_NAME = "date_groups"
# read in csv file
completeLines_df = read_csv(str_c(WD_DATA,csvFileName)) %>%
  mutate(!!TIME_ID_NAME := as.Date( !!sym(TIME_ID_NAME) ))
```

```{r}
colnames(completeLines_df)
```

## get list of wavelengths, lineIDs, and timeIDs

```{r}
# vec of lineIDs  in completeLines_df
lineIDs = completeLines_df %>% group_by(!!sym(LINE_ID_NAME)) %>% summarize(n. = n()) %>% pull(!!sym(LINE_ID_NAME))
# vec of timeIDs in completeLines_df
timeIDs = completeLines_df %>% group_by(!!sym(TIME_ID_NAME)) %>% summarize(n. = n()) %>% pull(!!sym(TIME_ID_NAME))
T_ = length(timeIDs)
L_ = length(lineIDs)
```


# summaries of lines

```{r}
completeLines_df %>%
  group_by(!!sym(LINE_ID_NAME), !!sym(TIMEGROUP_ID_NAME)) %>%
  summarize( rmse = sqrt(mean( (rv_template_0.5-mean(rv_template_0.5))^2)),
             cor_depth = cor(rv_template_0.5, fit_gauss_depth),
             cor_width = cor(rv_template_0.5, fit_gauss_sigmasq),
             cor_hg0 = cor(rv_template_0.5, proj_hg_coeff_0),
             cor_hg1 = cor(rv_template_0.5, proj_hg_coeff_1),
             cor_hg2 = cor(rv_template_0.5, proj_hg_coeff_2)) %>%
  arrange(!!sym(LINE_ID_NAME))
```

wavelength and order coverage and mean depth
```{r}
completeLines_df %>%
  group_by(line_order) %>%
  summarize(wavelength = lambdac_vald[1],
            order = order_phys[1],
            mean_depth = mean(fit_gauss_depth),
            snr = mean(mean_snr_pixel)) %>%
  summary()
```


table of species
```{r}
completeLines_df %>%
  group_by(species) %>%
  count(line_order) %>%
  ungroup() %>%
  count(species) %>%
  arrange(n)
```


# rv and depth visuals for sample of lineIDs (visual in paper)

```{r}
selectLines = c("4732.02805_44","5408.26971_61","6244.82532_76", "8250.39009_100")
```

```{r}
plot_df = completeLines_df %>%
  filter(line_order %in% selectLines) %>%
  group_by(line_order) %>%
  mutate(mean_depth = mean(fit_gauss_depth),
         change_depth = fit_gauss_depth - mean_depth ) %>%
  ungroup() %>%
  mutate(lineID = str_c("wavelength = ", round(lambdac_vald,3), "\nmean depth = ", round(mean_depth,3))) %>%
  select(date, date_groups, lineID, rv_template_0.5, fit_gauss_depth, change_depth)
```

```{r}
rv_time_plt = plot_df %>%
  ggplot() +
  geom_point(mapping = aes(x = date, y = rv_template_0.5, color = date), size = .8, alpha = .87) +
  xlab("Date") +
  ylab("RV (m/s)") +
  facet_wrap(vars(lineID), nrow = 1, scales = "fixed") +
  theme_bw() + 
  guides(color="none") +
  scale_color_viridis_c() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 20),
        strip.text.x = element_text(size = 15) )
depth_time_plt = plot_df %>%
  ggplot() +
  geom_point(mapping = aes(x = date, y = change_depth, color = date ), size = .8, alpha = .87) +
  guides(color="none") +
  xlab("Date") +
  ylab("Change in Depth") +
  facet_wrap(vars(lineID), nrow = 1, scales = "fixed") +
  theme_bw() +
  theme(strip.text.x = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 20)) + 
  guides(color="none") +
  scale_color_viridis_c()
rv_depth_plt = plot_df %>%
  ggplot() +
  geom_point(mapping = aes(x = change_depth, y = rv_template_0.5, color = date ), size = .8, alpha = .87) +
  guides(color="none") +
  xlab("Change in Depth") +
  ylab("RV (m/s)") +
  facet_wrap(vars(lineID), nrow = 1, scales = "free") +
  theme_bw() +
  theme(strip.text.x = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 20))+ 
  guides(color="none") +
  scale_color_viridis_c()

final_plot = rv_time_plt/depth_time_plt/rv_depth_plt

# print the final arrangement
print(final_plot)
ggsave("visuals/fourLinesRV.pdf", width = 13, height = 10)
``` 

```{r}
rm(selectLines, plot_df, rv_time_plt, depth_time_plt, rv_depth_plt, final_plot)
```

# heatmap, rv by date (visual in paper)

```{r}
# rv limit of plot
rvLim = 40

# data frame of the plot, get number of observations removed from plot
plotDF = completeLines_df %>%
  mutate(date = as.factor(date)) %>%
  filter(abs(rv_template_0.5) < rvLim)
cat(nrow(completeLines_df) - nrow(plotDF), "removed observations\n")
cat( 100*(nrow(completeLines_df) - nrow(plotDF))/nrow(completeLines_df), "percent removed\n" )
cat(L_, "total lineIDs\n")

# non-adjusted rv plot
allLinesHeat_plt = plotDF %>%
  ggplot() +
  geom_bin2d(mapping = aes(x = date, y = rv_template_0.5), bins = 90  ) +
  geom_vline(xintercept = "2022-12-06", alpha = .5, linewidth=1) +
  scale_fill_gradient(low="orange",high="blue", trans="log10",limits = c(1,100)) +
  labs(x = "",
       y = "Pre-adjusted RV (m/s)",
       fill = "# of lines") +
  ylim(-rvLim - 1, rvLim + 1) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title=element_text(size=20))

# get a df of the adjusted RV (ie centered by the date group and lineID)
plotDF_adjusted = completeLines_df %>%
  group_by(date_groups, line_order) %>%
  mutate(rv_template_0.5 = rv_template_0.5 - mean(rv_template_0.5)) %>%
  ungroup() %>%
  mutate(date = as.factor(date)) %>%
  filter(abs(rv_template_0.5) < rvLim)
cat(nrow(completeLines_df) - nrow(plotDF_adjusted), "removed observations\n")
cat( 100*(nrow(completeLines_df) - nrow(plotDF_adjusted))/nrow(completeLines_df), "percent removed\n" )
cat(L_, "total lineIDs\n")

allLinesHeat_plt_adjusted = plotDF_adjusted %>%
  ggplot() +
  geom_bin2d(mapping = aes(x = date, y = rv_template_0.5), bins = 90  ) +
  geom_vline(xintercept = "2022-12-06", alpha = .5, linewidth=1) +
  scale_fill_gradient(low="orange",high="blue", trans="log10",limits = c(1,100)) +
  labs(x = "Chronologically-Ordered Days",
       y = "Post-adjusted RV (m/s)",
       fill = "# of lines") +
  ylim(-rvLim - 1, rvLim + 1) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title=element_text(size=20))

final_plot = (allLinesHeat_plt / allLinesHeat_plt_adjusted) + plot_layout(guides = 'collect') &
  theme(legend.position='bottom', legend.title=element_text(size=15))
print(final_plot)

#ggsave(filename = "visuals/heatmapRV_combined.pdf", width = 10, height = 7, units = "in")

rm(allLinesHeat_plt, rvLim, plotDF, plotDF_adjusted,allLinesHeat_plt_adjusted,final_plot)
```

# visual of absorption lines

```{r}
library(patchwork)
```

```{r}
# function to generate a shifted/activity influenced absorption line
generate_absorptionLine = function(x, shift = 0, sigma = 1) {
  y <- 1 - dnorm(x-shift, sd = sigma)
  return(y)
}
# function to generate an "activity influenced absorption line" as mixture of normal
mixture_density = function(x, sds, weights, means, shift = 0) {
  # ensure the weights sum to 1
  weights = weights / sum(weights)
  # calculate the density of each component and take the weighted sum
  density = sapply(x, function(xi) {
    sum(weights * dnorm(xi - shift, mean = means, sd = sds))
  })
  
  return(1-density)
}
```

```{r}
# get initial wavelengths for reference line 
length_out = 400
left_thresh = -3.5
right_thresh = 3.5
ref_wavelengths = seq(left_thresh, right_thresh, length.out = length_out)
ref_fluxs = generate_absorptionLine(ref_wavelengths)
```


```{r}
# with planet, no activity
planetYesActivityNo = generate_absorptionLine(ref_wavelengths, shift = .3)
# with activity, no planet
planetNoActivityYes = mixture_density(ref_wavelengths, sds = c(1,.3), means = c(0,.15), weights = c(.95,.05), shift = 0)
# with activity and planet
planetYesActivityYes = mixture_density(ref_wavelengths, sds = c(1,.3), means = c(0,.15), weights = c(.95,.05), shift = .3)

# y limits
y_lims = c( min(planetYesActivityNo,planetNoActivityYes,planetYesActivityYes), 
            max(planetYesActivityNo,planetNoActivityYes,planetYesActivityYes))


plt1 = ggplot() +
  geom_line(mapping = aes(x = ref_wavelengths, y = ref_fluxs), color = "black", linewidth = .9, alpha = .8) +
  geom_line(mapping = aes(x = ref_wavelengths, y = planetYesActivityNo), color = "red", linewidth = .9, alpha = .8) +
  theme_classic() +
  theme(
    axis.title=element_text(size=25),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 25)
  ) +
  ggtitle("Absorption line\nwith a planet, no activity") +
  ylim(y_lims) +
  xlab("wavelength") +
  ylab("flux")
plt2 = ggplot() +
  geom_line(mapping = aes(x = ref_wavelengths, y = ref_fluxs), color = "black", linewidth = .9, alpha = .8) +
  geom_line(mapping = aes(x = ref_wavelengths, y = planetNoActivityYes), color = "red", linewidth = .9, alpha = .8) +
  theme_classic() +
  theme(
    axis.title=element_text(size=25),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 25)
  ) +
  ggtitle("Absorption line\nwith activity, no planet") + 
  ylim(y_lims) +
  xlab("wavelength") +
  ylab("")
plt3 = ggplot() +
  geom_line(mapping = aes(x = ref_wavelengths, y = ref_fluxs), color = "black", linewidth = .9, alpha = .8) +
  geom_line(mapping = aes(x = ref_wavelengths, y = planetYesActivityYes), color = "red", linewidth = .9, alpha = .8) +
  theme_classic() +
  theme(
    axis.title=element_text(size=25),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 25)
  ) +
  ggtitle("Absorption line\nwith a planet and activity") + 
  ylim(y_lims) +
  xlab("wavelength") +
  ylab("")

plt1+plt2+plt3
ggsave("visuals/example_absorptionLines.png", height = 4, width = 13.5)
```


```{r}
plt1 = ggplot() +
  geom_line(mapping = aes(x = ref_wavelengths, y = ref_fluxs), color = "black", linewidth = .9, alpha = .8) +
  geom_vline(xintercept =ref_wavelengths[which.min(ref_fluxs)], color = "black", linewidth = .9, alpha = .8, linetype = 2 ) +
  geom_line(mapping = aes(x = ref_wavelengths, y = planetYesActivityNo), color = "red", linewidth = .9, alpha = .8) +
    geom_vline(xintercept =ref_wavelengths[which.min(planetYesActivityNo)], color = "red", linewidth = .9, alpha = .8, linetype = 2 ) +
  theme_classic() +
  theme(
    axis.title=element_text(size=25),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 25)
  ) +
  ggtitle("") +
  ylim(y_lims) +
  xlab("wavelength") +
  ylab("flux")
plt1
ggsave("visuals/example_absorptionLine.png")
```

# example planet signal

```{r}
AMP = 3
HORIZONTAL_OFFSET = 0
VERTICAL_OFFSET = 0
PERIOD = 250
FREQ = 2*pi/PERIOD

seq_days = seq(0,1.5*PERIOD)
rv_signal = VERTICAL_OFFSET + AMP*sin(FREQ*changeFun(as.numeric(seq_days))+HORIZONTAL_OFFSET)
ggplot() +
  geom_line(mapping = aes(x = seq_days,y = rv_signal), linewidth = .6 ) +
  geom_hline(yintercept = 0, color = "gray", alpha = .7) +
  geom_segment( mapping = aes(x = PERIOD/4+15, y = AMP, xend = 5*PERIOD/4-15, yend = AMP),
                arrow = arrow(length = unit(0.03, "npc"), ends = "both") ) +
  annotate("text", x=190, y=AMP-.3, label= "Orbital period", size = 5) +
  geom_segment( mapping = aes(x = PERIOD/4, y = 0, xend = PERIOD/4, yend = AMP-.05),
                arrow = arrow(length = unit(0.03, "npc"), ends = "both") ) +
  annotate("text", x=PERIOD/4-20, y=AMP/2, label= "Amp", size = 5) +
  ylab("RV (m/s)") +
  xlab("Time (days)") +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 20))

ggsave("visuals/rv_example.png", width = 6.8, height = 3)

rm(AMP,HORIZONTAL_OFFSET,VERTICAL_OFFSET,FREQ,PERIOD,rv_signal,seq_days)
```


# pairs plots for lines


```{r}
# covariates to visaulize
covars = c("hg_coeff_0","fit_gauss_depth")
# (random) sample of lines
sampleLines = sample(lineIDs,4)

completeLines_df %>%
  filter(line_order %in% sampleLines) %>%
  ggplot() +
  geom_point(mapping = aes(x = !!sym(covars[1]), y = !!sym(covars[2]), color = factor(date_groups) ), size = .6) +
  guides(color="none") +
  facet_wrap(vars(line_order), scales = "free")

rm(sampleLines,covars)
```


# snr

SNR by wavelength/order
```{r}
completeLines_df %>%
  group_by(!!sym(LINE_ID_NAME)) %>%
  summarize(mean_snr = mean(mean_snr_pixel),
            sd_snr = sd(mean_snr_pixel),
            wavelength = mean(fit_gauss_lambdac),
            order_id = mean(order_idx)) %>%
  ggplot() +
  geom_point(mapping = aes(x = wavelength, y = mean_snr))
completeLines_df %>%
  group_by(!!sym(LINE_ID_NAME)) %>%
  summarize(mean_snr = mean(mean_snr_pixel),
            sd_snr = sd(mean_snr_pixel),
            wavelength = mean(fit_gauss_lambdac),
            order_id = mean(order_idx)) %>%
  ggplot() +
  geom_point(mapping = aes(x = order_id, y = mean_snr))
```


```{r}
100*completeLines_df %>%
  group_by(!!sym(LINE_ID_NAME)) %>%
  summarize(mean_snr = mean(mean_snr_pixel),
            sd_snr = sd(mean_snr_pixel),
            wavelength = mean(fit_gauss_lambdac),
            order_id = mean(order_idx)) %>%
  filter(mean_snr <= 300) %>%
  nrow() /778
```



