---
title: "NEID Solar Data -- Linear Modeling Results"
author: "Joe Salzer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("readSpectra.R")
library(Matrix)
library(parallel)
library(pbmcapply)
#library(lme4)
#library(car)
#library(emmeans)
#library(boot)
# autocorrelation
#library(collapse)
#library(glmm)
#library(MASS)
#library(gridExtra)
#library(glmnet)
#library(plm)
#library(sandwich)
#library(lmtest)
```

```{r}
# working directory with the data
wd_data = "/Users/josephsalzer/research/exostat/"
# RESPONSE variable
RESPONSE = "rv_template_0.5"
# names of the timeID, lineID, and timeGroupID
TIME_ID_NAME = "date"
LINE_ID_NAME = "line_order"
TIMEGROUP_ID_NAME = "date_groups"
```

```{r}
model_name = "Gauss=all_HG=all"
```

*Options for model name:*

TWFE (Baseline model)
Gauss=all_HG=none
Gauss=none_HG=all
Gauss=all_HG=all (Full model)
CommonSlopes
LASSO (Full Model w/ LASSO)

# fit results

```{r}
model_fit = readRDS(str_c(wd_data, "models/", model_name, "/model.rds" ))

# get the lm fit and dataframe
designMat = model_fit$designMat
responses = model_fit$responses
rv_df = model_fit$df
group_sizes = model_fit$group_sizes
modelFormula = model_fit$modelFormula
covariates = model_fit$covariates
fit_lm = model_fit$fit_lm
#leverages = readRDS(str_c(wd_data, "models/", model_name, "/leverages.rds" ))
RMSE = model_fit$RMSE
```

```{r}
# vec of LineIDs  in completeLines_df
lineIDs = rv_df %>% group_by(!!sym(LINE_ID_NAME)) %>% summarize(n. = n()) %>% pull(!!sym(LINE_ID_NAME))
# vec of timeIDs in completeLines_df
timeIDs = rv_df %>% group_by(!!sym(TIME_ID_NAME)) %>% summarize(n. = n()) %>% pull(!!sym(TIME_ID_NAME))
T_ = length(timeIDs)
L_ = length(lineIDs)
```

```{r}
# dataframe of coefficients of current model
coef_df = tibble( coef = rownames( fit_lm$beta_hat ),
                  estimate = fit_lm$beta_hat[,1],
                  se = sqrt( diag(fit_lm$var_beta_hat) ) )
```

```{r}
# clean RV df with +- se
cleanRV_df = model_fit$cleanRV_df %>%
  mutate(lowerSE_MLE = estimate - se,
         upperSE_MLE = estimate +se)
```

result summary
```{r}
cat(model_name, "\n",
    str_c("AIC: ", round(fit_lm$AIC/1e5,3), "e5" ), "\n",
    str_c("BIC: ", round(fit_lm$BIC/1e5,3), "e5" ), "\n",
    str_c("RSE: ", round(fit_lm$RSE,4), "\n" ),
    str_c("Parameters: ", ncol(designMat), "\n"),
    str_c("RMSE of cleaned RV to true planet: ", round(model_fit$RMSE,3) ), "\n",
    str_c("mean se: ", round(mean( cleanRV_df$se ),3) ), "\n",
    str_c("coverage 1se: ",
          round(mean((cleanRV_df$estimate - cleanRV_df$se < 0) & (cleanRV_df$estimate + cleanRV_df$se > 0)),3),
          "\n"),
    str_c("coverage 2se: ",
          round(mean((cleanRV_df$estimate - 2*cleanRV_df$se < 0) & (cleanRV_df$estimate + 2*cleanRV_df$se > 0)),3),
          "\n"))
```

results for table in paper

```{r}
cat(model_name, "&",
    ncol(designMat), "&",
    str_c( "$", round(fit_lm$AIC/1e5,3), " \\times 10^5$"), "&",
    str_c( "$", round(fit_lm$BIC/1e5,3), " \\times 10^5$"), "&",
    round(fit_lm$RSE,3), "&",
    round(model_fit$RMSE,3), "\\\\"
    )
```

BF
Gauss=all_HG=all & 11896 & $6.604 \times 10^5$ & $7.803 \times 10^5$ & 6.279 & 0.578 \\
AF
Gauss=all_HG=all & 11772 & $2.812 \times 10^5$ & $3.906 \times 10^5$ & 5.405 & 0.461 \\

## diagnostic plots

rv effect plot

```{r}
ggplot( mapping = aes(x = cleanRV_df$timeID) ) +
  geom_point(mapping = aes(y = cleanRV_df$estimate), size = 1.2) +
  geom_hline(yintercept = 0, color = "red") +
  ylab("Cleaned RV") +
  xlab("Date") +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 20))
```

fitted rv and residual plot
```{r}
ggplot() +
  geom_point(mapping = aes(x = fit_lm$y_hat[,1], y = fit_lm$resid[,1]), size = .15, alpha = .5) +
  theme_minimal() +
  labs(title = "fitted rv and residual plot",
       x = "fitted rv",
       y = "residuals") +
  ylim(-100,100) +
  xlim(-200,200)
```

QQ plot
```{r}
# QQ plot
ggplot(data.frame(sample = fit_lm$resid[,1]), aes(sample = sample)) +
  stat_qq(size = .5, alpha = .5) +
  stat_qq_line() +
  labs(title = "QQ Plot",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()
```

leverages
```{r}
ggplot() + geom_density(mapping = aes(x = leverages))
summary(leverages)
```

autocorrelation for a single line
```{r}
lineID = 10
acf(rv_df %>% mutate(residual = fit_lm$resid[,1]) %>% filter(line_order %in% lineIDs[lineID]) %>% pull(residual))
rm(lineID)
```

correlation between lineIDs

```{r}
# lineID = sample(1:length(lineIDs), 2, replace = F)
lineID = which(lineIDs %in% c("5347.28348_59","5502.99268_63"), arr.ind = T)

rv_df %>%
  mutate(residual = fit_lm$resid[,1]) %>%
  filter(line_order %in% lineIDs[lineID]) %>%
  select(line_order, date, residual) %>%
  pivot_wider(id_cols = date, names_from = line_order, values_from = residual) %>%
  rename(line1 = lineIDs[lineID[1]], line2 = lineIDs[lineID[2]] ) %>%
  ggplot() +
  geom_point(mapping = aes(x = line1, y = line2))
rm(lineID)
```
correlation matrix between line's residuals

```{r}
# resid_cor_mat = rv_df %>%
#   mutate(residual = fit_lm$resid[,1]) %>%
#   select(line_order, date, residual) %>%
#   pivot_wider(id_cols = date, names_from = line_order, values_from = residual) %>%
#   select(!date) %>%
#   cor()
# 
# # correlation threshold
# threshold = 0.3
# 
# # find the column pairs with correlation higher than the threshold
# high_corr_pairs = which(abs(resid_cor_mat) > threshold, arr.ind = TRUE)
# # Filter out the diagonal and lower triangular values to avoid duplicate pairs
# high_corr_pairs = high_corr_pairs[high_corr_pairs[, 1] < high_corr_pairs[, 2], ]
# 
# # Convert the results to a readable format
# results = data.frame(
#   Column1 = rownames(resid_cor_mat)[high_corr_pairs[, 1]],
#   Column2 = colnames(resid_cor_mat)[high_corr_pairs[, 2]],
#   Correlation = resid_cor_mat[high_corr_pairs]
# )
# 
# # Print the results
# print(results)
```


looking at the slope covariates of the lines, seeing if they are similar

```{r}
coef_df %>%
  filter(endsWith(coef,"fit_gauss_depth_centered")) %>%
  arrange(estimate)
```
```{r}
rv_df %>%
  filter(line_order=="4885.95763_48") %>%
  ggplot(mapping = aes(x = as.Date(date), y = fit_gauss_depth)) +
  geom_point()
rv_df %>%
  filter(line_order=="4885.95763_48") %>%
  ggplot(mapping = aes(x = as.Date(date), y = rv_template_0.5)) +
  geom_point()
rv_df %>%
  filter(line_order=="4885.95763_48") %>%
  ggplot(mapping = aes(x = fit_gauss_depth, y = rv_template_0.5)) +
  geom_point()
```


weights by variance of resids

```{r}
rv_df %>%
  mutate(residual = fit_lm$resid[,1]) %>%
  group_by(line_order) %>%
  summarize( weight_byLine = mean(1/(sigma_rv_template_0.5)^2),
             resid_sd = sd(residual)) %>%
  mutate( weight_byLine = L_/sum(weight_byLine) * weight_byLine) %>%
  ggplot() +
  geom_point(mapping = aes(x = weight_byLine, y = log(resid_sd) ))
```


## RMSE COMPARED TO AN INJECTED SIGNAL

Finding the optimal RMSE for injected planetary signals to ensure consistent results across any injected planet. The cleaned RV RMSE estimates don't change as we include or remove planets

```{r}
true_rv_df = rv_df %>%
  select(date,date_groups,pert_val) %>%
  unique() %>%
  mutate(alpha_hat = cleanRV_df$estimate) %>%
  rename(true_rv = pert_val) %>%
  group_by(date_groups) %>%
  mutate(true_rv_offsetGroup = true_rv - mean(true_rv)) %>%
  ungroup()
true_rv_df
# group offset
true_rv_df %>%
  group_by(date_groups) %>%
  summarize(dateGroup_offset = mean(true_rv))
```

comparing our alpha hats to the (offset) true signal produces the same rmse as the model without a planet:
```{r}
rmse_t(c = 0, v = true_rv_df$alpha_hat - true_rv_df$true_rv_offsetGroup)
```

```{r}
ggplot( mapping = aes(x = cleanRV_df$timeID) ) +
  geom_point(mapping = aes(y = cleanRV_df$estimate), size = .7) +
  geom_errorbar(mapping = aes(ymin = cleanRV_df$lowerSE_MLE, ymax = cleanRV_df$upperSE_MLE), color = "black",linewidth = .1) +
  geom_hline(yintercept = 0, color = "red") +
  # with planet
   #geom_line(mapping = aes(x = cleanRV_df$date[1] + seq(0,918), y = 0+5*sin( (2*pi/366)*seq(0,918) + 30)), color = "red") +
  ylab("clean rv") +
  xlab("day") #+ ylim(-6.1,6.1)
```


## figure of both TWFE and full model

```{r}
model_fit_TWFE = readRDS(str_c(wd_data, "models/", "TWFE", "/model.rds" ))
model_fit_FULL = readRDS(str_c(wd_data, "models/", "Gauss=all_HG=all", "/model.rds" ))

cleanRV_df = rbind(
  model_fit_TWFE$cleanRV_df %>% mutate(model_name = "Baseline model"),
  model_fit_FULL$cleanRV_df %>% mutate(model_name = "Full model")
)

rm(model_fit_TWFE,model_fit_FULL)

```

```{r}
cleanRV_df %>%
  ggplot(mapping = aes(x = timeID, y = estimate, color = model_name, shape = model_name)) +
  geom_point(size = 1.2, alpha = 0.85) +
  geom_hline(yintercept = 0, color = "black") +
  scale_color_manual(values = c("#ff7f00", "#377eb8")) +
  scale_shape_manual(values = c(16, 17)) +
  guides(
    color = guide_legend(title = "Model", override.aes = list(size = 4)),
    shape = guide_legend(title = "Model")
  ) +
  labs(
    y = "Cleaned RV - True Signal (m/s)",
    x = "Date"
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 20),
    legend.title = element_blank(),
    legend.text = element_text(size = 20),
    legend.position = "bottom"
  )

ggsave("visuals/cleanRV.pdf", width = 10, height = 5)
```

```{r}
cleanRV_df %>%
  ggplot(mapping = aes(x = timeID, y = estimate, color = model_name, shape = model_name)) +
  geom_point(size = 1, alpha = .85) +
  #geom_ribbon(aes(ymin = estimate-1, ymax = estimate+1), alpha = .3 ) +
  geom_hline(yintercept = 0, color = "black") +
  scale_color_manual(values=c("#ff7f00", "#377eb8")) +
  guides(color=guide_legend(title="Model",override.aes = list(size=5))) +
  ylab("Cleaned RV - True Signal (m/s)") +
  xlab("Date") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 20), 
        legend.title= element_blank(),
        #legend.title= element_text(size = 15),
        legend.text=element_text(size=20),
        legend.position="bottom")

ggsave("visuals/cleanRV.pdf", width = 10, height = 5)
```




# bootstrap results

```{r}
bootName = "/wild_bootstraps"

boot_files = list.files( str_c(wd_data, "models/", model_name, bootName) )
coef_list = list()

for ( i in 1:length(boot_files) ) {
  coef_boot = readRDS( str_c(wd_data, "models/", model_name, bootName, "/", boot_files[i] ) )
  coef_list[[i]] = coef_boot
}

# extract all vectors from boot_straps
coef_boot_df = do.call( rbind, coef_list )

rm(coef_boot, boot_files, coef_list, bootName)
```

```{r}
# get group sizes
group_sizes = rv_df %>%
  group_by(!!sym(TIMEGROUP_ID_NAME)) %>%
  summarize(size = n()/L_) %>%
  pull(size)

# initialize 0's in the linear operator
linear_op_mat = Matrix(0, nrow = T_, ncol = length(fit_lm$beta_hat[,1]), sparse = T )
# matrix for estimating the cleaned RV
linear_op_mat[,(length(group_sizes)+1):sum(group_sizes)] = bdiag(lapply(group_sizes, function(n) contr.sum(n) ))
```

```{r}
coef_boot_df[1:5,1:5]
```

```{r}
# get the cleaned RV, after applying the linear operator to get the clean RV estimates for each day
# each column is the estimate for each day and each row is a boot sample
cleanRV_boot_df = coef_boot_df %*% t(linear_op_mat)
```

```{r}
dim(coef_boot_df)
dim(linear_op_mat)
dim(cleanRV_boot_df)
```

```{r}
coef_boot_df[1:5,1:5]
cleanRV_boot_df[1:5,1:5]
```

```{r}
# check for bias in original estimate vs bootstrap estimate
id = 10

ggplot() +
  geom_histogram(mapping = aes(x = cleanRV_boot_df[,id]), bins = 40) +
  geom_vline(xintercept = cleanRV_df$estimate[id] )

rm(id)
```

```{r}
# add the bootstrap se to the cleanRV_df
cleanRV_df$boot_se = apply(cleanRV_boot_df, 2, sd)

# ci based on se of sampling dist
cleanRV_df = cleanRV_df %>%
    mutate(lowerSE_boot = estimate - boot_se,
           upperSE_boot = estimate + boot_se) 
```

```{r}
# overall rmse for each boot sample
rmse_boot = apply(cleanRV_boot_df, 1, rmse_t, c = 0)

ggplot() +
  geom_density(mapping = aes(x = rmse_boot)) +
  geom_vline(xintercept = model_fit$RMSE, linetype = 2) +
  xlab("RMSE")

summary(rmse_boot)

model_fit$RMSE
quantile(rmse_boot, probs = c(.005,.995))
```



```{r}
cleanRV_df %>%
  ggplot() +
  geom_point(mapping = aes(x = timeID, y = boot_se), size = 1.2, color = "black", alpha = .8) +
  geom_hline(yintercept = mean(cleanRV_df$se), color = "red", linetype = 2, linewidth = 1.5) +
  ylab("Standard Error of Cleaned RV") +
  xlab("Date") +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 23))

ggsave("visuals/se_cleanRV.pdf", width = 10, height = 5)
```



```{r}
summary(cleanRV_df$se)
```

*coverage comparisons*
find how much we have to multiply the se by to get a certain coverage

```{r}
findMult4Coverage = function(estimate, se, desired_coverage, tol = .0001) {
  mult = 1
  current_coverage = mean((estimate - mult*se < 0) & (estimate + mult*se > 0))
  while (current_coverage < desired_coverage) {
    mult = mult + tol
    current_coverage = mean((estimate - mult*se < 0) & (estimate + mult*se > 0))
  }
  # CI length at desired level of coverage
  CI_lengths = 2*mult*se
  
  return(c(mult = mult,
           current_coverage = current_coverage,
           mean_CI = mean(CI_lengths),
           med_CI = median(CI_lengths),
           quant95 = unname(quantile(CI_lengths,.95)),
           quant05 = unname(quantile(CI_lengths,.05)),
           quant25 = unname(quantile(CI_lengths,.25)),
           quant75 = unname(quantile(CI_lengths,.75))
           ))
}
```

MLE coverage
```{r}
N = 250
mleCoverage = seq(from = .68,to = 1,length.out = N)
mleCoverage_df = data.frame( matrix(NA,ncol=9) )
colnames(mleCoverage_df) = c("desired_coverage","mult","current_coverage","mean_CI","med_CI","quant95","quant05","quant25","quant75")

for (i in 1:N) {
  mleCoverage_df[i,] = c(mleCoverage[i],findMult4Coverage(cleanRV_df$estimate,cleanRV_df$se,mleCoverage[i]))
  if (i %% 10 == 0) print(i)
}

rm(N,i,mleCoverage)
```

boot coverage

```{r}
N = 250
bootCoverage = seq(from = .68,to = 1,length.out = N)
bootCoverage_df = data.frame( matrix(NA,ncol=9) )
colnames(bootCoverage_df) = c("desired_coverage","mult","current_coverage","mean_CI","med_CI","quant95","quant05","quant25","quant75")

for (i in 1:N) {
  bootCoverage_df[i,] = c(bootCoverage[i],
                          findMult4Coverage(cleanRV_df$estimate,cleanRV_df$boot_se,bootCoverage[i]))
  if (i %% 10 == 0) print(i)
}

rm(N,i,bootCoverage)
```

```{r}
mleCoverage_df = mleCoverage_df %>%
  mutate(uncertainty_method = "MLE")
bootCoverage_df = bootCoverage_df %>%
  mutate(uncertainty_method = "boot")
coverage_df = rbind(mleCoverage_df,bootCoverage_df)
```

```{r}
coverage_df  %>%
  ggplot() +
  geom_line(mapping = aes(x = desired_coverage, y = mult, color = uncertainty_method))
```

```{r}
ggplot() +
  geom_line(mapping = aes(x = coverage_df$desired_coverage,
                          y = coverage_df$med_CI,
                          color = coverage_df$uncertainty_method),
            linewidth = 1) +
  geom_ribbon(mapping = aes(x = bootCoverage_df$desired_coverage,
                            ymin = bootCoverage_df$quant05,
                            ymax = bootCoverage_df$quant95),
              fill="gray", alpha=0.5) +
  theme_bw() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 20),
        legend.text=element_text(size=20),
        legend.position="bottom") +
  scale_y_log10() +
  xlab("Coverage") +
  ylab("Confidence Interval Length (m/s)") +
  scale_colour_manual(labels = c("Bootstrap","MLE"),
                      values = c("black","red")) +
  guides(color=guide_legend(override.aes = list(linewidth=2)))

ggsave("visuals/CI_coverage.pdf", width = 7, height = 7)
```


*using quantile-based coverage for bootstrap*


```{r}
cleanRV_boot_df[1:5,1:5]
dim(cleanRV_boot_df)
```

Each row is a bootstrap sample and each column is the bootstrap clean RV for a day. I tried to evaluate quantiles (at various levels) of each bootstrap clean RV to determine the bootstrap confidence interval for each of the cleaned RVs.

```{r}
alpha_ = 0
upper_quant = apply(cleanRV_boot_df, 2, quantile, probs = 1-alpha_/2)
lower_quant = apply(cleanRV_boot_df, 2, quantile, probs = alpha_/2)
mean( (upper_quant > 0) & (lower_quant < 0) )
summary(upper_quant-lower_quant)
```
```{r}
upper_quant = apply(cleanRV_boot_df, 2, max)
lower_quant = apply(cleanRV_boot_df, 2, min)
mean( (upper_quant > 0) & (lower_quant < 0) )

rm(alpha_, upper_quant, lower_quant)
```

unfortunately, with the above we will never reach a coverage above 0.869697, because the max - min only ever achieves a coverage of 0.869697

```{r}
findQuant4Coverage = function(boot_samples, desired_coverage, tol = .001) {
  alpha_ = 0
  upper_quant = apply(cleanRV_boot_df, 2, quantile, probs = 1-alpha_/2)
  lower_quant = apply(cleanRV_boot_df, 2, quantile, probs = alpha_/2)
  current_coverage = mean( (upper_quant > 0) & (lower_quant < 0) )
  
  while (current_coverage > desired_coverage) {
    alpha_ = alpha_ + tol
    upper_quant = apply(cleanRV_boot_df, 2, quantile, probs = 1-alpha_/2)
    lower_quant = apply(cleanRV_boot_df, 2, quantile, probs = alpha_/2)
    current_coverage = mean( (upper_quant > 0) & (lower_quant < 0) )
  }
  
  # CI length at desired level of coverage
  CI_lengths = upper_quant-lower_quant
  
  return(c(alpha_ = alpha_,
           current_coverage = current_coverage,
           mean_CI = mean(CI_lengths),
           med_CI = median(CI_lengths),
           quant95 = unname(quantile(CI_lengths,.95)),
           quant05 = unname(quantile(CI_lengths,.05)),
           quant25 = unname(quantile(CI_lengths,.25)),
           quant75 = unname(quantile(CI_lengths,.75))
           ))
}
```

```{r}
N = 100
bootCoverage = seq(from = .68,to = 0.86,length.out = N)

bootQuantCoverage_df = data.frame( matrix(NA,ncol=9) )
colnames(bootQuantCoverage_df) = c("desired_coverage","mult","current_coverage","mean_CI","med_CI","quant95","quant05","quant25","quant75")

for (i in 1:N) {
  bootQuantCoverage_df[i,] = c(bootCoverage[i],
                          findQuant4Coverage(cleanRV_boot_df,bootCoverage[i]))
  if (i %% 10 == 0) print(i)
}
```


```{r}
bootQuantCoverage_df
```

```{r}
bootQuantCoverage_df = bootQuantCoverage_df %>%
  mutate(uncertainty_method = "quantile_boot")

coverage_df = rbind(coverage_df,bootQuantCoverage_df)
```

median ci length by coverage (by method)
```{r}
coverage_df %>%
  filter(desired_coverage < 0.84) %>%
  ggplot() +
  geom_line(mapping = aes(x = desired_coverage,
                          y = med_CI,
                          color = uncertainty_method),
            linewidth = 1) +
  theme_bw() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 12),
        legend.text=element_text(size=12),
        legend.position="bottom") +
  scale_y_log10() +
  xlab("Coverage") +
  ylab("Confidence Interval Length") +
  guides(color=guide_legend(override.aes = list(linewidth=2)))

```

```{r}
coverage_df %>%
  filter(desired_coverage < 0.84) %>%
  ggplot() +
  geom_line(mapping = aes(x = desired_coverage,
                          y = med_CI,
                          color = uncertainty_method),
            linewidth = 1) +
  geom_line(mapping = aes(x = desired_coverage,
                          y = quant25,
                          color = uncertainty_method),
            linewidth = 1,
            linetype = 3) +
    geom_line(mapping = aes(x = desired_coverage,
                          y = quant75,
                          color = uncertainty_method),
            linewidth = 1,
            linetype = 3) +
  theme_bw() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 12),
        legend.text=element_text(size=12),
        legend.position="bottom") +
  xlab("Coverage") +
  ylab("Confidence Interval Length") +
  guides(color=guide_legend(override.aes = list(linewidth=2)))
```



# cross validation

```{r}
cvNames = c("LOODCV","LOOWCV","LOOMCV")
modelNames = c("TWFE","Gauss=all_HG=all")

# empty list for creating data frame
cv_df = data.frame()

for (cvName in cvNames) {
  for (modelName in modelNames) {
    # list of files
    cv_files = list.files( str_c(wd_data, "models/", modelName, "/", cvName) )
    # empty list for creating data frame
    cv_list = list()
    # time ID
    timeIDs = rep(NA,length(cv_files))
    # initialize cleaned RV's and standard errors 
    rv_decontam = rep(NA, length(timeIDs))
    se_decontam = rep(NA, length(timeIDs))
  
    for ( i in 1:length(cv_files) ) {
      # day that we'll test on
      test_day = str_sub(str_split_fixed(cv_files, "_",3)[i,3],end = -5)
      timeIDs[i] = test_day
      
      # the entire test set df
      testset_df = readRDS( str_c(wd_data, "models/", modelName, "/", cvName, "/", cv_files[i] ) )$testDF
      # grab the test day from the validation set 
      decontamRV = testset_df %>%
        filter(date == test_day) %>%
        mutate(cleaned_rv = contam_rv - pred_rv) %>%
        pull(cleaned_rv)
      
      # get the cleaned RV and se
      rv_decontam[i] = mean(decontamRV)
      se_decontam[i] = sd(decontamRV)

      
      rm(test_day,testset_df)
    }
    
    cv_df = rbind(cv_df,
                  data.frame(date = timeIDs,
                             decontamRV = rv_decontam,
                             seRV = se_decontam,
                             modelID = modelName,
                             cv_procedure = cvName
                             )
                  )
    
    #print(tail(cv_df))

  }
}
```






```{r}
cvNames = c("LOODCV","LOOWCV","LOOMCV")
modelNames = c("TWFE","LASSO","Gauss=all_HG=all")

# empty list for creating data frame
cv_df = data.frame()

for (cvName in cvNames) {
  for (modelName in modelNames) {
    # list of files
    cv_files = list.files( str_c(wd_data, "models/", modelName, "/", cvName) )
    # empty list for creating data frame
    cv_list = list()
    # time ID
    timeIDs = rep(NA,length(cv_files))
    # initialize cleaned RV's and standard errors 
    rv_decontam = rep(NA, length(timeIDs))
    se_decontam = rep(NA, length(timeIDs))
  
    for ( i in 1:length(cv_files) ) {
      # day that we'll test on
      test_day = str_sub(str_split_fixed(cv_files, "_",3)[i,3],end = -5)
      timeIDs[i] = test_day
      
      # the entire test set df
      testset_df = readRDS( str_c(wd_data, "models/", modelName, "/", cvName, "/", cv_files[i] ) )$testDF
      # grab the test day from the validation set 
      decontamRV = testset_df %>%
        filter(date == test_day) %>%
        mutate(cleaned_rv = contam_rv - pred_rv) %>%
        pull(cleaned_rv)
      
      # get the cleaned RV and se
      rv_decontam[i] = mean(decontamRV)
      se_decontam[i] = sd(decontamRV)

      
      rm(test_day,testset_df)
    }
    
    cv_df = rbind(cv_df,
                  data.frame(date = timeIDs,
                             decontamRV = rv_decontam,
                             seRV = se_decontam,
                             modelID = modelName,
                             cv_procedure = cvName
                             )
                  )
    
    #print(tail(cv_df))

  }
}
```

rmse results by model
```{r}
cv_df %>%
  group_by(modelID,cv_procedure) %>%
  summarize(rmse = rmse_t(0,decontamRV),
            rmse_optim = rmse_t(0,decontamRV-mean(decontamRV)))
```

```{r}
cv_df %>%
  filter(modelID != "LASSO",
         cv_procedure == "LOOMCV") %>%
  mutate(modelID = ifelse(modelID == "TWFE","Baseline model","Full Model") ) %>%
  ggplot() +
  geom_point(mapping = aes(x = as.Date(date), y = seRV, color = modelID, shape = modelID), size = 1.4, alpha = .85) +
  theme_minimal() +
  scale_y_log10() +
  scale_color_manual(values=c("#ff7f00", "#377eb8")) +
  scale_shape_manual(values = c(16, 17)) +
  guides(
    color = guide_legend(title = "Model", override.aes = list(size = 5)),
    shape = guide_legend(title = "Model")
  ) +
  ylab("Standard Error (m/s)") +
  xlab("Date") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 20), 
        legend.title= element_blank(),
        #legend.title= element_text(size = 15),
        legend.text=element_text(size=20),
        legend.position="bottom")

ggsave("visuals/cv_se.pdf", width = 10, height = 5)
```

```{r}
cv_df %>%
  filter(modelID != "LASSO",
         cv_procedure == "LOOMCV") %>%
  mutate(modelID = ifelse(modelID == "TWFE","BaselineModel","FullModel") ) %>%
  pivot_wider(id_cols = date, names_from = modelID, values_from = seRV) %>%
  mutate(percentDiff = 100*(BaselineModel-FullModel)/FullModel ) %>%
  summarize(avgPercentDiff = mean(percentDiff))
cv_df %>%
  filter(modelID != "LASSO",
         cv_procedure == "LOOWCV") %>%
  mutate(modelID = ifelse(modelID == "TWFE","BaselineModel","FullModel") ) %>%
  pivot_wider(id_cols = date, names_from = modelID, values_from = seRV) %>%
  mutate(percentDiff = 100*(BaselineModel-FullModel)/FullModel ) %>%
  summarize(avgPercentDiff = mean(percentDiff))
cv_df %>%
  filter(modelID != "LASSO",
         cv_procedure == "LOODCV") %>%
  mutate(modelID = ifelse(modelID == "TWFE","BaselineModel","FullModel") ) %>%
  pivot_wider(id_cols = date, names_from = modelID, values_from = seRV) %>%
  mutate(percentDiff = 100*(BaselineModel-FullModel)/FullModel ) %>%
  summarize(avgPercentDiff = mean(percentDiff))
```

```{r}
cv_df %>%
  filter(modelID != "TWFE",
         cv_procedure == "LOODCV") %>%
  pivot_wider(id_cols = date, names_from = modelID, values_from = seRV) %>%
  mutate(isLassoBetter = LASSO<=!!sym("Gauss=all_HG=all")) %>%
  summarize(percentLASSO = mean(isLassoBetter),
            mean_diff = mean((LASSO-!!sym("Gauss=all_HG=all"))^2))
cv_df %>%
  filter(modelID != "TWFE",
         cv_procedure == "LOOWCV") %>%
  pivot_wider(id_cols = date, names_from = modelID, values_from = seRV) %>%
  mutate(isLassoBetter = LASSO<=!!sym("Gauss=all_HG=all")) %>%
  summarize(percentLASSO = mean(isLassoBetter),
            mean_diff = mean((LASSO-!!sym("Gauss=all_HG=all"))^2))
cv_df %>%
  filter(modelID != "TWFE",
         cv_procedure == "LOOMCV") %>%
  pivot_wider(id_cols = date, names_from = modelID, values_from = seRV) %>%
  mutate(isLassoBetter = LASSO<=!!sym("Gauss=all_HG=all")) %>%
  summarize(percentLASSO = mean(isLassoBetter),
            mean_diff = mean((LASSO-!!sym("Gauss=all_HG=all"))^2))
```

LASSO is slightly better for LOOMCV, slightly worse for LOODCV, LOOWCV.


```{r}
cv_df %>%
  filter(modelID != "LASSO",
         cv_procedure == "LOOMCV") %>%
  ggplot() +
  geom_point(mapping = aes(x = as.Date(date), y = decontamRV, color = modelID), size = 1) +
  theme_minimal() +
  facet_wrap(~cv_procedure)
```